Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
minor nit fixes
edits to 5

\paragraph*{Open and Closed Regular Expressions} While \GreedySynth acts
relatively efficiently, it suffers from an exponential blowup when converting
regular expressions to DNF form. This can be mitigated by not converting
\emph{all} of the regular expression to DNF form, but only converting
\emph{some} of it, and doing so incrementally. This change is easy
to support by adding non-DNF stochastic regular expressions as a new type of
``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
annotated as ``closed'', then $\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
Regular expressions that are not annotated as closed, are considered ``open.''

At the start of synthesis, all regular expressions are closed. \Expand is
modified so that it is now permitted to rewrite closed regular expressions to
open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
map closed regular expressions to each other, but only by using the identity
lens---they must parse the same string components in all examples.

\paragraph*{Expansion Inference} Distinguishing open and closed regular
expressions helps \GreedySynth, but doing so means that \Expand has the harder job of
determining when to open closed expressions.  To that end,
expansion inference identifies when a regular expression cannot
possibly map by any means other than the ``trivial'' disconnect lens. Expansion inference
opens those regular expressions to allow for further exploration.

\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
memoizing its results.  However, our stochastic DNF regular expressions are
quite large and so hashing them is computationally intensive.  We use less
memory and allow for quick comparisons and hashing of these DNF regular
expressions by hash-consing them.

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses---whether they are previously defined by users or by
synthesis. We alter \GreedySynth such that two regular expression atoms can be
mapped to each other either if they are the same regular expression and agree
with the examples, or if there is an existing lens (which is a composition of predefined
lenses) that already agrees with the examples.

\paragraph*{Termination Customization}
In this work, there are three competing constraints. (1) We want to return
lenses quickly to users, (2) we want to return good lenses, and (3) we don't
want to require users to write a large number of examples in their
specifications. Our termination heuristic works well for most situations, but
there are cases where a large number of expansions must be performed to find the
correct lens.  Users could work around that problem by providing a large number
of examples, but it can be difficult to write a large number of informative
examples for complex formats. We find it simpler to instead let the users
specify that \SynthSymLens should be more restrictive in which lenses it
accepts.  Conversely, if the user's desired transformation has a particularly
high cost, the algorithm may spend a long time searching in vain for lenses with
lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
short-circuited.

For these situations, users can provide a number, $t$, which affects
the strictness of the termination condition, adjusting the \Continue formula
as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
\end{gather*}
With a positive $t$, the termination condition is stricter, and the algorithm
will spend longer searching for more likely lenses. With a negative $t$ the
termination condition is less strict, and the algorithm will spend less time
searching for more likely lenses.  Our default $t$ is $0$.

While our defaults termination condition quickly finds the correct lens for
39 of our 48 benchmarks, our termination condition is not
one-size-fits all. Complex lenses, or lenses that project a
large amount of data, can be synthesized more easily and quickly by
adjusting $t$.
% end synthesis

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used 11 benchmarks from synchronization tasks between a variety of
  formats taken from the bidirectional programming literature. These tasks range
  from synchronizing REST and JSON web resource descriptions to synchronizing
  BibTeX and EndNote citation descriptions.
\end{enumerate}

@@

\subsection{Useful in Everyday Development}

>>>>

\paragraph*{Open and Closed Regular Expressions} While \GreedySynth acts
relatively efficiently, it can suffer from an exponential blowup when converting
regular expressions to DNF form. This problem can be mitigated by avoiding the conversion
of some REs to DNF form and by performing the conversion lazily when necessarily.
More specifically, \Expand labels some uncoverted REs as ``closed,'' which means the
type-directed \GreedySynth algorithm treats them as DNF atoms and does not dig into
them recursively.  In other words,
given a pair of closed REs, \GreedySynth can either construct
the identity lens between them (it will do this if they are the same RE), or it can
construct a disconnect lens between them.
Regular expressions that are not annotated as closed are considered ``open.''
%is allowed to to
%map closed regular expressions to each other, but only by using the identity
%lens---they must parse the same string components in all examples.
%This change is easy
%to support by adding non-DNF stochastic regular expressions as a new type of
%``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
%annotated as ``closed'', then $\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.

\paragraph*{Expansion Inference} Distinguishing between open and closed regular
expressions improves the efficiency of \GreedySynth, but forces \Expand to decide
which closed expressions to open.
At the start of synthesis, all regular expressions are closed. \Expand 
rewrites selected closed regular expressions to
open ones (thereby triggering DNF normalization)
when the disconnect lens is the only lens that will ever apply.

\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
hash-consing stochastic DNF regular expressions.  This optimization enables
fast equality checks and saves memory.\dpw{we could probably cut this paragraph.}
%memoizing its results.  However, our stochastic DNF regular expressions are
%quite large and so hashing them is computationally intensive.  We use less
%memory and allow for quick comparisons and hashing of these DNF regular
%expressions by

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses, whether they are previously defined by users or by
synthesis. As the synthesizer processes a Boomerang file, it accumulates
lens definitions and their types.  It tackles synthesis problems one after another
and there may be many such problems in a given program.
During synthesis, if an existing lens has the right type and agrees with the examples.
\GreedySynth will use it.  Consequently, if a large synthesis problem cannot be solved all
at once, a user can first generate subproblems for parts of a data format, and the
larger problem can subsequently use solutions to those subproblems.  In our experience,
this is a very powerful tool that allows synthesis to tackle problems of arbitrary
complexity.

\paragraph*{Termination Customization}
In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where a large number of expansions must be performed to find the
correct lens.  Users can work around such a problem by providing a large number
of examples, but constructing examples is a chore, especially when formats are
complicated.  We have often found it simpler to allow users to change the
synthesis termination condition to demand \SynthSymLens be more judgemental
when it comes to deciding it has found a ``good enough'' lens.  
Conversely, if the user's desired transformation has a particularly
high cost, the algorithm may spend a long time searching in vain for lenses with
lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
short-circuited.

For these situations, users can provide a number, $t$, that affects
the strictness of the termination condition, adjusting the \Continue formula
as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
\end{gather*}
With a positive $t$, the termination condition is stricter, and the algorithm
will spend longer searching for more likely lenses. With a negative $t$, the
termination condition is less strict, and the algorithm will spend less time
searching for more likely lenses.  Our default $t$ is $0$.

As discussed in the following section,
while our default termination condition quickly finds the correct lens for
39 of our 48 benchmarks, our termination condition is not
one-size-fits all. Complex lenses, or lenses that project a
large amount of data, can be synthesized more easily and quickly by
adjusting $t$. \dpw{should this paragraph be cut and its content left for the
next section?}
% end synthesis

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used 11 benchmarks from synchronization tasks between a variety of
  formats derived from real-world examples and the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions.
\end{enumerate}

@@

\subsection{Usefulness in Everyday Development}
